---
title: "Assignment 1"
author: 'Group 47: Luca Cavellini, Paul Hosek, Robert Satzger'
date: "2/20/2023"
output:
  pdf_document: default
  html_document: default
fontsize: 11pt
highlight: tango
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE, message = FALSE, fig.height = 3)

# setting standard R plot themes to be used by ggplots
library(ggplot2)
theme_set(theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  panel.background = element_blank(), 
  axis.line = element_line(colour = "black"))) 
```

# Preparation: Defining Re-Usable Assumption Checks
To standardize assumption checks throughout the assignment, we use the following functions. `check_normality` plots the histogram, qq-plot, and boxplots for all variables of interest and returns the shapiro-wilk test. `check_vars` extends assumption checks for ANOVAs by showing the qq-plot of the residuals and plotting the fitted values against the residuals.

```{r, code = readLines("../utilities.R")}
```

```{r child = 'w1ex1.Rmd'}
```

```{r child = 'w1ex2.Rmd'}
```

```{r child = 'w1ex3.Rmd'}
```

```{r child = 'w1ex4.Rmd'}
```