---
title: "ex2_cholesterol"
author: "Paul Hosek, .......add more names here"
date: "2023-02-17"
output: pdf_document
---

# Exercise 2. Cholesterol

```{r, include=FALSE}
source("../utilities.R")
```

```{r}
library(ggplot2)
library(dplyr)
library(ggpubr)
data <- read.table("data/cholesterol.txt", header = TRUE, sep = " ")
```

**a)**

First plotting the discrete distribution of values, we see that values appear 
approximately normally distributed. Further, QQ-plots indicate that samples 
from both measurements are similarly distributed by inspection.
However, we need to confirm this with thorough statistical testing. 

```{r, echo=TRUE, fig.height = 4}
shapiros <- check_normality(list("Before", "After8weeks"), data)
```

Shapiro-Wilk tests for normality indicate that the assumption of normality
is met for both pre- and post-measurement. P-values are $p = `r sprintf('%.2f',shapiros[[1]]$p.value)`$ for the 
before measurement and  $p = `r sprintf('%.2f',shapiros[[2]]$p.value)`$ for the post measurement respectively.


Further, we observe a very high correlation of the data. The Pearson correlation coefficient is $r = 0.99$ with significant correlation of pre and post measurement $p = 0.99$.
```{r, echo=TRUE, fig.height = 2}

ggscatter(data, x = colnames(data)[1], y = colnames(data)[2],
          add = "reg.line",cor.coef = TRUE, conf.int = TRUE,
          cor.method = "pearson",
          xlab = "Before", ylab = "After 8 Weeks");

```



**b)**
Since both measurements are taken from the same experimental units, we perform two paired samples tests. Based on the assumption of normality in **a**, we perform a dependent, two-samples t-test. For comparison, we also perform a non-parametric alternative: The Mann-Whitney rank-based test.

```{r, echo=TRUE}
ttest <- t.test(data$Before, data$After8weeks, paired = TRUE, alternative = "two.sided");
print(ttest)

wilcoxtest <- wilcox.test(data$Before, data$After8weeks, 
                          paired = TRUE, alternative = "two.sided");
print(wilcoxtest)
```

Both the t-test and the Mann-Whitney test concur in that there is a significant effect of the diet on weight, with $p = `r sprintf('%.2f',ttest$p.value)`$ and $p = `r sprintf('%.2f',wilcoxtest$p.value)`$, respectively. However, the statistical effect does not imply a large effect size. Therefore, the practical utility of the diet is not necessarily given.

The permutation test is applicable, since we have a test statistic that expresses a difference between dependent measures. While the assumption of normality is met in this case, the permutation test would also apply in case of a violation of this assumption.

**c)**

Based on the Central Limit Theorem, we know that the test statistic of any distribution will be normally distributed as we resample.

```{r, echo=TRUE}
nsamples <- 1000;

theta <- replicate(nsamples, max(sample(data$After8weeks,size=18, replace = TRUE)));
Tstar25=quantile(theta,0.025)
Tstar975=quantile(theta,0.975)
T1 = max(data$After8weeks)
CI = c(2*T1-Tstar975,2*T1-Tstar25)
print(CI)
```

The confidence interval is $CI = [`r sprintf('%.2f',CI[[1]])`,`r sprintf('%.2f',CI[[2]])`]$.

This confidence interval could be improved by resampling more often.
However, since the sample size is only 18, conclusions about the true population mean are limited. As such, we recommend to collect more data.

**d)**
```{r, echo=TRUE, fig.height=2.5}
nsamples <- 1000;
thetas <- seq(3, 12, by = .01);
t <- max(data$After8weeks)
p_vals <- numeric(length(thetas));
for (i in 1:length(thetas)) {
  res <- replicate(nsamples,max(runif(18,3, thetas[i])))
  pl=sum(res<t)/nsamples
  pr=sum(res>t)/nsamples
  p_vals[i]=2*min(pl,pr)
}

plot(thetas, p_vals, type = "l", xlab = "theta", ylab = "p-value");
abline(h = 0.05, col = "red");
legend("topright", legend = c("p-value", "0.05"),
       col = c("black", "red"), lty = c(1, 1));
t <- thetas[p_vals >.05]
I <- c(min(t), max(t))
print(I)
```
The thetas for which the hypothesis is not rejected lie in the interval $I = [`r sprintf('%.2f', I[[1]])`,`r sprintf('%.2f',I[[2]])`]$.

Alternatively, we can use the Kolmogorov-Smirnov test to test if 
the distributions are the same. But we must be careful to not accept
distributions with a maximum value smaller than the maximum of the sample, because else the maximum lies outside the given range.

```{r, echo=TRUE, console=FALSE, warning = F}
thetas <- seq(3, 12, by = .01);

# Loop over theta values and perform bootstrap test
ks_res <- numeric();
p_vals <- numeric(length(thetas));
for (i in 1:length(thetas)) {
  theta <- thetas[i];
  cur_unif <- runif(10000,3,thetas[i]);
  p_vals[i] <- ks.test(data$After8weeks, cur_unif)['p.value'];
}

```

```{r, echo=FALSE, console=FALSE}
plot(thetas, p_vals, type = "l", xlab = "theta", ylab = "p-value");
abline(h = 0.05, col = "red");
legend("topright", legend = c("p-value", "0.05"),
       col = c("black", "red"), lty = c(1, 1));

```
**e)**

```{r, echo=TRUE}
binom_e <- binom.test(sum(data$After8weeks < 6), nrow(data), alternative = "less")
```

The binomial test returns an insignificant p-value of $p = `r sprintf('%.2f', binom_e$p.value)`$. Therefore, we cannot reject the null hypothesis stating that the median post measurement is less than 6.

First, we find true fraction of cholesterol levels less than 4.5. 

```{r, echo=TRUE}
binom_e1 <- binom.test(sum(data$After8weeks < 4.5), nrow(data), 
                       p = 0.25, alternative = "l")
print(binom_e1)
binom_e2 <- binom.test(sum(data$After8weeks < 4.5), nrow(data), 
                       p = 0.2500001, alternative = "l")
print(binom_e2)
```

With the null hypothesis H0: p >= 0.25, the binomial test returns an insignificant p-value of $p = `r sprintf('%.2f', binom_e1$p.value)`$. However, the alternative hypothesis should include the value 0.25. Therefore, we corroborated our findings with a second test, including the value 0.25. Therefore, we can not confirm the claim that the fraction of cholesterol levels less than 4.5 is at most 25%. 