---
title: "ex2_cholesterol"
author: "Paul Hosek, .......add more names here"
date: "2023-02-17"
output: pdf_document
---

# Exercise 2. Cholesterol

```{r, include=FALSE}
source("../utilities.R")
```

```{r}
library(ggplot2)
library(dplyr)
library(ggpubr)
data <- read.table("data/cholesterol.txt", header = TRUE, sep = " ")
```

**a)**

First plotting the discrete distribution of values, we see that values appear 
approximately normally distributed. Further, QQ-plots indicate that samples 
from both measurements are similarly distributed by inspection.
However, we need to confirm this with thorough statistical testing. 

```{r, echo=TRUE, fig.height = 4}
shapiros <- check_normality(list("Before", "After8weeks"), data)
# data_long <- data.frame(
#   Time = rep(c("Before", "After 8 Weeks"), each = nrow(data)),
#   Value = c(data$Before, data$After8weeks)
# );
# ggplot(data_long, aes(x = Value, fill = Time)) +
#   geom_histogram(alpha = 0.5, position = "identity", bins = 10) +
#   labs(title = "Histogram of Before and After 8 Weeks", x = "Values", y = "Frequency") +
#   scale_fill_manual(values = c("#E69F00", "#56B4E9"), labels = c("Before", "After 8 Weeks"));
```

Shapiro-Wilk tests for normality indicate that the assumption of normality
is met for both pre- and post-measurement. P-values are $p = `r sprintf('%.2f',shapiros[[1]]$p.value)`$ for the 
before measurement and  $p = `r sprintf('%.2f',shapiros[[2]]$p.value)`$ for the post measurement respectively.


Further, we observe a very high correlation of the data.
```{r, echo=TRUE, fig.height = 2.5}

ggscatter(data, x = colnames(data)[1], y = colnames(data)[2],
          add = "reg.line",cor.coef = TRUE, conf.int = TRUE,
          cor.method = "pearson",
          xlab = "Before", ylab = "After 8 Weeks");

```

**b)**
First, we use a non-parametric, repeated measures t-test.
```{r, echo=TRUE}
t.test(data$Before, data$After8weeks, paired = TRUE, alternative = "two.sided");
```
```{r, echo=TRUE}
wilcox.test(data$Before, data$After8weeks, paired = TRUE, alternative = "two.sided");
```

The results indicate, that at $\alpha =0.05$, there is a significant effect
of the diet. However, this must not implyeffect has practical significance and
is strong enough for the diet to be useful in practice.

The permutation test is applicable, since it can express any test-statistic
including repeated measures and also is non-parameteric.

**c)**
```{r, echo=TRUE}
nsamples <- 1000;

theta <- replicate(nsamples, max(sample(data$After8weeks,size=18, replace = TRUE)));
cat("The confidence interval of the max is:",quantile(theta, c(0.025, 0.975)),"\n");
```
This confidence interval could be improved by resampling more often.
However, since the sample size is only 18, conclusions about the true population
mean are limited. As such, we recommend to collect more samples.

**d)**
```{r, echo=TRUE}
nsamples <- 1000;
thetas <- seq(3, 12, by = .01);
t <- max(data$After8weeks)
p_vals <- numeric(length(thetas));
for (i in 1:length(thetas)) {
  res <- replicate(nsamples,max(runif(18,3, thetas[i])))
  pl=sum(res<t)/nsamples
  pr=sum(res>t)/nsamples
  p_vals[i]=2*min(pl,pr)
}

plot(thetas, p_vals, type = "l", xlab = "theta", ylab = "p-value");
abline(h = 0.05, col = "red");
legend("topright", legend = c("p-value", "0.05"),
       col = c("black", "red"), lty = c(1, 1));
```
Alternatively, we can use the Kolmogorov-Smirnov test to test if 
the distributions are the same. But we must be careful to not accept
distributions with a maximum value smaller than the maximum of the sample,
because else the maximum lies outside the given range.
```{r, echo=TRUE, console=FALSE}
thetas <- seq(3, 12, by = .01);

# Loop over theta values and perform bootstrap test
ks_res <- numeric();
p_vals <- numeric(length(thetas));
for (i in 1:length(thetas)) {
  theta <- thetas[i];
  cur_unif <- runif(10000,3,thetas[i]);
  p_vals[i] <- ks.test(data$After8weeks, cur_unif)['p.value'];
}

```

```{r, echo=FALSE, console=FALSE}
plot(thetas, p_vals, type = "l", xlab = "theta", ylab = "p-value");
abline(h = 0.05, col = "red");
legend("topright", legend = c("p-value", "0.05"),
       col = c("black", "red"), lty = c(1, 1));

```
**e)**

```{r, echo=TRUE}
binom_e <- binom.test(sum(data$After8weeks < 6), nrow(data), alternative = "less")
cat("The p-value does not reject the H0 at alpha=0.05: p-value =",binom_e$p.value)
```
First, we find true fraction of cholesterol levels less than 4.5. 
```{r, echo=TRUE}
prop <- mean(data$After8weeks < 4.5)
cat("The fraction of cholesterol levels <4.5 is", prop, "..")

```

```{r, echo=TRUE}
binom_e <- binom.test(sum(data$After8weeks < 4.5), nrow(data), p = 0.25, alternative = "less")
cat("No, the fraction <4.5 is not less than 25%, with p=",binom_e$p.value)
```