---
title: "w2ex3"
author: "Paul Hosek"
date: "`r Sys.Date()`"
output: pdf_document
---


```{r, include=FALSE}
source("../utilities.R")
library(ggplot2)
library(dplyr)
library(ggpubr)
library(car)
library(knitr)
```

- models will only work with !nan ages

#3
## a)
- multiple summaries of data
- fit log regression,, w/o interactions -> survival & predictor Pclass, age, sex
```{r, echo =TRUE}
data_titanic <- read.table("titanic.txt", header=TRUE)
data_titanic$PClass <- as.factor(data_titanic$PClass)
data_titanic$Sex <- as.factor(data_titanic$Sex)
data_titanic$Survived <- as.factor(data_titanic$Survived)
```


```{r, echo =TRUE}
par(mfrow=c(1,3))

ggplot(subset(data_titanic, Sex == "male"),
       aes(x = Age, fill = factor(Survived))) +geom_histogram(binwidth = 5)+scale_fill_discrete(labels = c("Not survived", "Survived"))+labs(title = "Survival by age: Males")

ggplot(subset(data_titanic, Sex == "female"), aes(x = Age, fill = factor(Survived))) +
  geom_histogram(binwidth = 5)+scale_fill_discrete(labels = c("Not survived", "Survived"))+labs(title = "Survival by age: Females")

ggplot(data_titanic, aes(x = PClass, fill = factor(Survived))) +
  geom_bar()+scale_fill_discrete(labels = c("Not survived", "Survived"))+labs(title = "Survival by class")
```
```{r, echo =TRUE}
model_log1 <- glm(Survived ~ PClass + Age + Sex, data = data_titanic, family = binomial())
summary(model_log1)
```
Excluding interaction effects, we find that being a female or a first class passengers or young increases your odds of survival. However, we cannot know how a combination of these will impact the odds. From the main effects we can conclude: Males are $`r sprintf('%.2f',1/exp(coef(model_log1)["Sexmale"]))`$ more likely to die compared to females. 2nd-class passengers are $`r sprintf('%.2f',1/exp(coef(model_log1)["PClass2nd"]))`$ and third-class passengers are $`r sprintf('%.2f',1/exp(coef(model_log1)["PClass3rd"]))`$ as likely to die than passengers in other classes (calculated as 1/exp(coefficient_of_interest)). Further, for each year a person is older, odds decrease by a factor of $`r sprintf('%.2f',exp(coef(model_log1)["Age"]))`$: younger passengers are more likely to survive (calculated as exp(age)). All these main effects are statistically significantly associated with survival.
## b)
```{r, echo =TRUE}
model_log2 <- glm(Survived ~ PClass + Age + Sex + PClass:Age + Age:Sex, data = data_titanic, family = binomial)
summary(model_log2)
```

```{r, echo =TRUE}
all_comb_55 <- expand.grid(PClass = levels(data_titanic$PClass), Sex = levels(data_titanic$Sex), Age = 55)
all_comb_55$Survival_Probability <- predict(model_log2, all_comb_55, type = "response") # response = probabilities, "link" for odds
kable(all_comb_55, format = "latex", caption = "Survival probability for 55 year olds.")
```
We observe that being female has the largest influence on survival. Independent of gender, more expensive classes have larger survival probability. We observe that females in the first class have a extremely high survival probability of $`r sprintf('%.2f',all_comb_55$Survival_Probability[[1]])`$.
```{r, echo =TRUE}
```
## c)

- method to predict
  - logistic regression
  - MCMC
  - machine learning
  - idk bro lecture is not out yet
- measure of quaility
  - validation set
  - area under ROC
```{r, echo =TRUE}
```

```{r, echo =TRUE}
```
