---
title: "w2ex2.Rmd"
author: "Robert Satzger"
date: "3/9/2023"
output: html_document
---

# Exercise 1. Trees
**a)**

First, we test the normality assumption of the ANOVA.

```{r}
trees = read.table("data/treeVolume.txt", header = T)
trees$type <- as.factor(trees$type)
qqnorm(trees$volume); qqline(trees$volume)
```

The QQ-plot suggests that the data may deviate from the normality assumption. However, the ANOVA is generally robust against such violations when there are more than ten observations per condition. Therefore, we will proceed with the analysis.

```{r}
trees_lm1 <- lm(volume ~ type, data = trees)
summary(trees_lm1)
trees_aov <- anova(trees_lm1)
print(trees_aov)
```

The one-way ANOVA returned a p-value of $p = 0.17$. Therefore, we cannot refute the null hypothesis that there is no influence of tree type on volume. It is noteworthy that a one-way ANOVA is generally endorsed in situations with a factor of more than two levels. In this case, the factor has two levels only. Thus, the result is equal to the result of a t-test (shown in the summary of the linear model). Based on our results, the estimated volume for the beech trees is 30.17, while for the oak trees it is 35.25.

**b)**
```{r}
trees_lm2 <- lm(volume ~ type * diameter + height, data = trees)
anova(trees_lm2)
trees_lm3 <- lm(volume ~ type * height + diameter, data = trees)
anova(trees_lm3)
```
In order to check whether diameter and height have the same influence on the volume for both tree types, we fit two models. The p-values returned from our tests do not support an interaction effect of tree type with either diameter or height on volume. Therefore, we can conclude that the effects of both height and diameter on volume are similar for both tree types.

**c)**
Based on the results in **b**, we continue with an exclusively additive model.

```{r}
trees_lm4 <- lm(volume ~ diameter + height + type, data = trees)
summary(trees_lm4)
drop1(trees_lm4, test='F')
```

The results suggest that tree type is not a significant factor in this model. Thus, we further reduce the linear model to only contain the predictors diameter and height.

```{r}
tree_lm5 <- lm(volume ~ diameter + height, data = trees)
smry <- summary(tree_lm5)
smry
```



```{r}
smry$coefficient["(Intercept)","Estimate"]+
  smry$coefficient["diameter","Estimate"]*mean(trees$diameter)+
  smry$coefficient["height","Estimate"]*mean(trees$height)
```

To predict the volume of a tree based on the diameter and the height we followed the formula 
$\hat{Y} = \hat{\mu} + \hat{\beta}_{diameter} \bar{X}_{diameter} + \hat{\beta}_{height} \bar{X}_{height}$, by which we obtained a value of 32.58.

**d)**
A cylinder's volume is given by $V = \pi * (d/2)^2 * h$. So, the most logical explanatory variable would be to calculate the volume using this formula.

```{r}
trees$transform <- pi * (trees$diameter/2)^2 * trees$height
trees_lm5 <- lm(volume ~ transform, data = trees)
summary(trees_lm5)

plot(trees$height, trees$diameter)
```

The transformation increases R-squared from .949 to .975. This shows the utility of using the correct mathematical formula in creating a new explanatory variable. It is noteworthy, however, that the fitness may be further improved by scaling the original predictors according to their units. As we are unaware of the units, we proceeded with the unscaled measures.
