---
title: "w2ex2.Rmd"
author: "Robert Satzger"
date: "3/9/2023"
output: html_document
---

# Exercise 1. Trees
**a)**
```{r}
trees = read.table("data/treeVolume.txt", header = T)
trees$type <- as.factor(trees$type)
trees_lm1 <- lm(volume ~ type, data = trees)
summary(trees_lm1)
trees_aov <- anova(trees_lm1)
print(trees_aov)
```
One-way Anova only makes sense with three or more levels. With two levels, p-value is the same as from t-test for difference of means between the two levels.
Based on summary of lm, beech is 30.17 and oak is 5.08.

**b)**
```{r}
trees_lm2 <- lm(volume ~ type * diameter + height, data = trees)
summary(trees_lm2)
anova(trees_lm2)
trees_lm3 <- lm(volume ~ type * height + diameter, data = trees)
summary(trees_lm3)
anova(trees_lm3)
```
Neither interaction effect seems to be present. Instead, use purely additive model.

```{r}
trees_lm4 <- lm(volume ~ type + diameter + height, data = trees)
summary(trees_lm4)
anova(trees_lm4)
```
<!-- predict overall average -->

**d)**
A cylinder's volume is given by $V = \pi * (d/2)^2 * h$. So most logical explanatory factor would be to calculate the volume using this formula.

<!-- What are the units? Normalize? -->
```{r}
trees$transform <- pi * (trees$diameter/2)^2 * trees$height
trees_lm5 <- lm(volume ~ transform, data = trees)
summary(trees_lm5)

plot(trees$height, trees$diameter)
```

R^2 shows better than the other models.